---
title: "BC Child Care Vacancies"
date: today
format:
  dashboard:
    theme: 
      - custom2.scss
    nav-buttons: 
      - icon: github
        href: https://github.com/wvictor14/bcchildcarebot
      - icon: house-fill
        href: https://victoryuan.com/
    include-in-header:
        - text: |
            <script src="reactable_row_details.js"></script>
editor_options: 
  chunk_output_type: console
execute:
  echo: false
---

```{r load_libraries}
#| include: FALSE
library(tidyverse)
library(glue)
library(reactable)
library(htmltools)
library(crosstalk)
library(leaflet)
library(tippy)
library(bslib)
library(here)
library(jsonlite)
source(here::here('R', 'filter.R'))
source(here::here('R', 'prepare_data.R'))
source(here::here('R', 'create_reactable.R'))
```

```{r load_data}
#| include: false
# read in bc child care data ----
url <- 'https://catalogue.data.gov.bc.ca/dataset/4cc207cc-ff03-44f8-8c5f-415af5224646/resource/9a9f14e1-03ea-4a11-936a-6e77b15eeb39/download/childcare_locations.csv'
bccc <- readr::read_csv(url)
.today <- lubridate::today(tzone = 'Canada/Pacific')
```

```{r setup_tbl_data}
# latest vacancy update
dataset_last_update <- max(bccc$VACANCY_LAST_UPDATE, na.rm = TRUE)

tbl_data <- bccc |>
  prepare_data() |> 
  
  # filter to last vacancy update - 14 days
  filter(VACANCY_LAST_UPDATE >= dataset_last_update - 1 -3) |> 
  arrange(desc(VACANCY_LAST_UPDATE))
shared_data <- SharedData$new(tbl_data)
```

# Main

## Filter Facilities {height=15%}

```{r filters}
div(
  class = "filters",
  search_filter("filter_name", "Search by Facility Name", shared_data, "NAME"),
  #select_filter("filter_city", "City", shared_data, "CITY"),
  crosstalk::filter_select('filter_city', 'City', shared_data, ~CITY),
  #select_filter("filter_vacancy", "Vacancy", shared_data, "VACANCY")
  filter_select2("filter_vacancy", "Vacancy", shared_data, ~VACANCY)
)
```

## Row 

```{r render_reactable}
#| padding: 0px
#| eval: true
# columns to hide from unested view
coldefs <- create_coldefs(tbl_data)
reactable(
  shared_data,
  defaultSortOrder = "desc",
  defaultColDef = colDef(
    # show = FALSE,
    headerClass = "header", align = "left"),
  onClick = "expand",
  resizable = TRUE,
  columns = coldefs,
  #details = function(index) row_details(index, tbl_data),
  details = colDef(
    details = JS('renderRowDetails'),
    html = TRUE,
  ),
  pagination = FALSE,
  searchable = FALSE,
  theme = reactableTheme(
    searchInputStyle = list(
      width = '100%'
    )
  )
)
```

```{r leaftlet_map}
#| column: screen
#| message: false
#| padding: 0px
icons <- awesomeIcons(
  icon = 'home',
  iconColor = 'black',
  library = 'glyphicon'
)

service_levels <- bccc$SERVICE_TYPE_CD |>  unique() |> sort()
pal <- colorFactor(
  c("navy", "firebrick", "orange"), 
  domain = service_levels
)

map <- shared_data |> 
  
  # map
  leaflet() |>
  setView(-123.04591, 49.25, zoom = 11) |> 
  
  addProviderTiles(providers$CartoDB.Voyager) |>
  #providers$CartoDB.Positron
  addCircleMarkers(
    popup = ~popup, 
    label = ~NAME,
    #usterOptions = markerClusterOptions(),
    #icon = icons
    stroke = FALSE,  fillOpacity = 0.6,
    color = ~pal(SERVICE_TYPE_CD),
    radius = 5
  ) |> 
  addLegend(
    "bottomright", pal = pal, 
    values = ~SERVICE_TYPE_CD,
    title = "Childcare type",
    opacity = 1
  )
map
```

# Info

Last date that this page has been updated: `{r} .today`

Last date that BC updated dataset: `{r} dataset_last_update`

Every day, new vacancies across BC Child Care facilities will be pull from the BC Childcare Map dataet with up-to-date vacancy information[^1].

[^1]: This dataset is not guaranteed to have accurate information. Please verify any information from this resource.

Data is downloaded every day from the BC government [BC Data Catalogue](https://catalogue.data.gov.bc.ca/dataset/child-care-map-data/resource/9a9f14e1-03ea-4a11-936a-6e77b15eeb39). However, the government does not necessarily update the dataset every day. This data also powers the [Child Care Map](https://maps.gov.bc.ca/ess/hm/ccf/), which the BC government maintains.
