---
title: "BC Child Care Vacancies"
date: today
format: 
  html:
    theme: 
      - custom2.scss
    nav-buttons: 
      - icon: github
        href: https://github.com/wvictor14/bcchildcarebot
      - icon: house-fill
        href: https://victoryuan.com/
editor_options: 
  chunk_output_type: console
execute:
  echo: false
---

```{r load_libraries}
#| include: FALSE
library(tidyverse)
library(glue)
library(reactable)
library(htmltools)
library(crosstalk)
library(leaflet)
library(tippy)
library(bslib)
library(here)
library(jsonlite)
source(here::here('R', 'filter.R'))
source(here::here('R', 'prepare_data.R'))
source(here::here('R', 'create_reactable.R'))
```

```{r load_data}
#| include: false
# read in bc child care data ----
url <- 'https://catalogue.data.gov.bc.ca/dataset/4cc207cc-ff03-44f8-8c5f-415af5224646/resource/9a9f14e1-03ea-4a11-936a-6e77b15eeb39/download/childcare_locations.csv'
bccc <- readr::read_csv(url)
.today <- lubridate::today(tzone = 'Canada/Pacific')
```

```{r setup_tbl_data}
# latest vacancy update
dataset_last_update <- max(bccc$VACANCY_LAST_UPDATE, na.rm = TRUE)

tbl_data <- bccc |>
  prepare_data() |> 
  
  # filter to last vacancy update - 14 days
  filter(VACANCY_LAST_UPDATE >= dataset_last_update - 1 -3) 
shared_data <- SharedData$new(tbl_data, key = ~id)
```

# Info

Last date that this page has been updated: `{r} .today`

Last date that BC updated dataset: `{r} dataset_last_update`

Every day, new vacancies across BC Child Care facilities will be pull from the BC Childcare Map dataet with up-to-date vacancy information[^1].

[^1]: This dataset is not guaranteed to have accurate information. Please verify any information from this resource.

# Filter Facilities

```{r filters}
div(
  class = "filters",
  search_filter("filter_name", "Search by Facility Name", shared_data, "NAME"),
  select_filter("filter_vacancy", "Vacancy", shared_data, "VACANCY"),
  range_filter("filter_last_updated", "Last Updated", shared_data, "VACANCY_LAST_UPDATE")
)
```

# Map

```{r}
#| column: screen
icons <- awesomeIcons(
  icon = 'home',
  iconColor = 'black',
  library = 'glyphicon'
)

service_levels <- bccc$SERVICE_TYPE_CD |>  unique() |> sort()
pal <- colorFactor(
  c("navy", "firebrick", "orange"), 
  domain = service_levels
)

map <- shared_data |> 
  
  # map
  leaflet() |>
  setView(-123.04591, 49.25, zoom = 11) |> 
  
  addProviderTiles(providers$CartoDB.Voyager) |>
  #providers$CartoDB.Positron
  addCircleMarkers(
    popup = ~popup, 
    label = ~NAME,
    #usterOptions = markerClusterOptions(),
    #icon = icons
    stroke = FALSE,  fillOpacity = 0.6,
    color = ~pal(SERVICE_TYPE_CD),
    radius = 5
  ) |> 
  addLegend(
    "bottomright", pal = pal, 
    values = ~SERVICE_TYPE_CD,
    title = "Childcare type",
    opacity = 1
  )
map
```

# Table 

```{r render_reactable}
#| padding: 0px
#| eval: true
# columns to hide from unested view
coldefs <- create_coldefs()
reactable(
  shared_data,
  #tbl_data[1:10,],
  defaultSorted = "VACANCY_LAST_UPDATE",
  defaultSortOrder = "desc",
  defaultColDef = colDef(
    # show = FALSE,
    headerClass = "header", align = "left"),
  onClick = "expand",
  resizable = TRUE,
  columns = coldefs,
  details = function(index) row_details(index, tbl_data),
  pagination = FALSE,
  searchable = TRUE,
  theme = reactableTheme(
    searchInputStyle = list(
      width = '100%'
    )
  )
) |> 
  bslib::card(
    height = '600px',
    full_screen = TRUE,
    style = "resize:vertical;"
)
```

# Info

Data is downloaded every day from the BC government [BC Data Catalogue](https://catalogue.data.gov.bc.ca/dataset/child-care-map-data/resource/9a9f14e1-03ea-4a11-936a-6e77b15eeb39). However, the government does not necessarily update the dataset every day. This data also powers the [Child Care Map](https://maps.gov.bc.ca/ess/hm/ccf/), which the BC government maintains.

It is well known that childcare is crazy in Vancouver. Personally, we experienced so many challenges in finding good childcare facilities with vacancies in Vancouver. So I built this childcare facility tool to help myself find childcare for my son. In my experience using this app we were able to very quickly find a nice childcare facility near our home that had vacancies. Maybe when he ages out into the next age group, I will update the app with more features.

I also used this experience to develop my professional development through learning how to build and deploy static html dashboards using quarto and github actions. It was a fun project and I learned several new tools.